### 2023-07-14 python_excel_ite.tex
first draft almost done - see TODOs

Two different approaches are shown to consolidate data from multiple sheets into a single sheet. One approach uses pandas while the other relies on a library specifically created for working with Excel files from Python, openpyxl \cite{bibid}. One advantage to using Python rather than Visual Basic for Applications for this task is that Python is well suited for working with data in many different formats including csv, json, xml and a myriad of database formats. The Excel file may be just one data source out of a collection of data sources that are used to create some sort of combined database. The generality of Python is a strength in such scenarios.

@@

Reshaping data \cite{bibid} from wide to long is a common data wrangling task especially in preparation for plotting. In this example, not only do students have to grapple with reshaping the data in an Excel workbook, they also have to deal with the all too common problem of multi-row header lines in a spreadsheet. Excel's flexibility makes it easy to create such multi-row headers and pity the poor analyst who then has to deal with it when trying to move the data into some sort of data frame or database table. This example is posed as a challenge to the students. Examples of the desired outputs are shown and a few hints given. The answer is provided at the end of the notebook. This example really shows the power of Python in creating a reproducible and automated approach to dealing with a poorly structured Excel workbook, creating not only restructured data frames but also non-trivial plots. Attempting to do all of this in Excel, even with VBA, is not an easy task.

@@

Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{bibid}. While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \cite{bibid}. Given that business students are very familiar with both Excel and Powerpoint, inclusion of this material in a business school Python based analytics course seems quite appropriate. Students can learn more advanced and extremely useful Python concepts and techniques within a context with which they are intimately familiar. Manipulation of MS Office directly relevant to the job content of many of the students. This teaching module is a natural follow up to the commonly taught spreadsheet based analytics courses found in many business schools and provides an interesting way to teach and learn Python for analytics.

\section{Random notes}
Learning good Python programming practices via learning to model in similar way that students learn good spreadsheet practices through building spreadsheet models using the W\&A textbook.

%https://www.bleepingcomputer.com/news/microsoft/microsoft-considers-adding-python-as-an-official-scripting-language-to-excel/

%https://www.reddit.com/r/excel/comments/lrw21h/will_ms_excel_ever_use_python_as_scripting/

>>>>

Two different approaches are shown to consolidate data from multiple sheets into a single sheet. One approach uses pandas while the other relies on a library specifically created for working with Excel files from Python, \href{https://openpyxl.readthedocs.io/en/stable/}{openpyxl}. One advantage to using Python rather than Visual Basic for Applications for this task is that Python is well suited for working with data in many different formats including csv, json, xml and a myriad of database formats. The Excel file may be just one data source out of a collection of data sources that are used to create some sort of combined database. The generality of Python is a strength in such scenarios.

@@

Reshaping, or tidying, data \cite{wickhamTidyData2014} from wide to long format is a common data wrangling task especially in preparation for plotting. In this example, not only do students have to grapple with reshaping the data in an Excel workbook, they also have to deal with the all too common problem of multi-row header lines in a spreadsheet. Excel's flexibility makes it easy to create such multi-row headers and pity the poor analyst who then has to deal with it when trying to move the data into some sort of data frame or database table. This example is posed as a challenge to the students. Examples of the desired outputs are shown and a few hints given. The answer is provided at the end of the notebook. This example really shows the power of Python in creating a reproducible and automated approach to dealing with a poorly structured Excel workbook, creating not only restructured data frames but also non-trivial plots. Attempting to do all of this in Excel, even with VBA, is not an easy task.

@@

Rumors have swirled for a number of years that Microsoft was considering adding Python as a first class language alternative to VBA within their MS Office suite of packages \cite{MicrosoftConsidersAdding}. While this has not yet materialized, there continues to be significant interest in using Python to manipulate not only Excel workbooks but other office documents such as Powerpoint presentations \cite{PythonpptxPythonpptx21}. Given that the ubiquity of both Excel and Powerpoint in business education and practice, inclusion of this material in a business school Python based analytics course seems quite appropriate. Students can learn more advanced and extremely useful Python concepts and techniques within a context with which they are intimately familiar. Manipulation of MS Office documents is directly relevant to the job content of many of the students. This teaching module is a natural complement to the commonly taught spreadsheet based analytics and introductory data science courses found in many business schools and provides an interesting way to teach and learn Python for analytics.
